### 2018-05-18 manuscript.rmd
incorporated author comments, fixed typos, updated references

Siphonophores remain poorly known, in large part because they are fragile and difficult to collect. They have, however, been of great interest for more than 150 years due to their unique structure and development [@mapstone2014global;@mackie1987siphonophore]. Like many other cnidarians, they are colonial: they grow by incomplete asexual reproduction. Each colony arises from a single embryo that forms the protozooid, the first body. One or two growth zones (Fig. \@ref(colony)) then arise and asexually produce other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1993ordre]. These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution]. Siphonophores differ significantly from other colonial animals in their colony structure and development -- their zooids are highly functionally specialized and arranged in precise, repeating, species-specific patterns [@Beklemishev1969;@cartwright2010character]. Zooids are specialized for a range of functions, including feeding, reproduction, or swimming (Fig. \@ref(colony)) [@dunn2006evolution].

@@

Understanding the unique ecology, morphology, and development of siphonophores requires a well-resolved phylogeny of the group. The relationship of siphonophores to other hydrozoans has been difficult to determine [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic; @kayal2015phylogenetic; @kayal2018phylogenomics], but there has been progress on their internal relationships. A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa addressed several long standing questions about siphonophore biology, including the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae. Cystonectae was found to be sister to all other siphonophores, while Calycophorae were nested within "Physonectae". The name Codonophora was given to this clade of "Physonectae" and Calycophorae [@Dunn:2005dy].

Major questions remained after this early work, though. In particular, there was little support for important deep relationships within Codonophora. Understanding these relationships is key to resolving the evolution of several traits of importance, including sexual systems (monoecy versus dioecy) and the gain and loss of particular zooids, such as palpons (Fig. \@ref(colony)). Here we present a broadly sampled phylogenetic analysis of Siphonophora that considers transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded due to poor sampling). Using 1,423 genes, we find strong support for many relationships found in the earlier phylogeny [@Dunn:2005dy], and also provide new resolution for key relationships that were unresolved in that previous study. Using this phylogeny, we reconstruct the evolutionary history of characters central to the unique biology of siphonophores, including zooid type, life history traits, and habitat.

@@

All scripts for the analyses are available in a git repository at [https://github.com/caseywdunn/siphonophore_phylogeny_2017](https://github.com/caseywdunn/siphonophore_phylogeny_2017). The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")` with tag "paper_v1".

@@

We used ModelFinder [@kalyaanamoorthy2017modelfinder], as implemented in IQTree v1.6.3 [@doi:10.1093/molbev/msu300], to assess relative model fit. ModelFinder selected XX based on XX. To assess the robustness of our results, we conducted phylogenetic analyses using multiple software programs, methods (Maximum likelihood (ML) and Bayesian Inference (BI)), and models (including the model selected by ModelFinder and several other commonly used models). Maximum likelihood (ML) analyses were conducted with RAxML v8.2.0 [@stamatakis2006raxml] and IQTree v1.6.3 [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281]. Bayesian Inference (BI) were conducted with Phylobayes v. 1.7a-mpi [@lartillot2009phylobayes]. Sequence alignments, sampled and consensus trees, and voucher information are available in the git repository. Tree figures were rendered with ggtree [@Yu:2016fo].

RAxML ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution (Figure XX). RAxML bootstrap values were estimated using `r length(bootstrap_trees)` replicates. IQTree ML analyses were run under JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories, the best model identified by ModelFinder, and the commonly used model GTR + Optimized base frequencies + Free rate model with 6 categories (Figure XX).

BI was conducted in phylobayes using two different models: fixed-state WAG+$\Gamma$ (Fig. S12) and CAT-Poisson (Figure S11). Eight chains were run under the CAT-Poisson model. The WAG+$\Gamma$ runs (XX provide the same detail as is available for CAT poisson). The CAT-Poisson runs did not converge (maxdiff=1, meandiff=0.0185032). Closer inspection revealed that chain1 and chain3 were stuck in local maxima with low likelihood relative to other chains after 1405 and 4695 iterations. These two chains were excluded from the analyses, and the results pressented here are based on the remaining 6 CAT-Poisson chains. Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs. This left `r length( phylobayes_trees )` trees in the CAT-Poisson posterior.

@@

Morphological character data used in trait mapping were obtained from the literature or direct observation of available voucher material. Depth distribution data was queried from the MBARI VARS database ( http://www.mbari.org/products/research-software/video-annotation-and-reference-system-vars/) [@schlining2006mbari]. We used stochastic character mapping to infer the probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169; @huelsenbeck2003stochastic]. Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package. See Supplementary Information for R package version numbers.

@@

The phylogenetic relationships recovered in this study received strong support across analysis methods (Fig. \@ref(phylogeny)A), with a couple of localized exceptions (Fig. \@ref(phylogeny)B and \@ref(bayesian)). All of the ML analyses were congruent with each other, regardless of model and software used (Fig. \@ref(fig:supplementary-trees)). These ML results were also congruent with the Phylobayes BI WAG+$\Gamma$ analyses (Fig. \@ref(fig:BI-WAG)). The Phylobayes BI CAT-Poisson result (Fig. \@ref(bayesian)), however, had a strongly supported topology that differed (Fig. \@ref(phylogeny)B) from the ML topology in a couple localized regions as described below. The fact that the Phylobayes BI WAG+$\Gamma$ is consistent with the WAG (and other) ML analyses suggests that the different topology recovered in the Phylobayes BI CAT-Poisson analyses is due to the different model rather than different software or methods. Here we take the conservative approach of considering relationships that differ between the Phylobayes BI CAT-Poisson analyses and other analyses to be unresolved.

@@

In ML analyses and BI WAG analyses, Euphysonectae consists of two reciprocally monophyletic groups that we here provisionally refer to as Clade A and Group B (Fig. \@ref(phylogeny)A). In BI CAT-Poisson analyses, Group B is paraphyletic (Fig. \@ref(phylogeny)B). The presence of an involucrum, a fold around the base of the cnidoband [@totton1965synopsis], is a potential synapomorphy for Clade A. Species of Clade A also have a descending mantle canal within the nectophores (Fig. \@ref(fig:desc-mantle), \@ref(fig:mantle-BI)), a structure that is also present in some calycophorans. Members of Clade A are also monoecious (Fig. \@ref(reconstruct)). There is not a clear synapomorphy for Group B. Within Group B there is high support for the placement of *Erenna richardi* in ML analyses (XX and BI WAG??), but it is placed as sister to Clade A in BI CAT-Poisson analyses (Fig. \@ref(phylogeny)B). More taxon sampling will be required to determine the relationship of species within this group.

@@

Nectophores (Fig. \@ref(colony)) are non-reproductive propulsive medusae. In Codonophora, the nectophores are localized to a region known as the nectosome (Fig. \@ref(colony)B), which has its own growth zone, and they are used for coordinated colony-level swimming. Planktonic cystonects like *Bathyphysa sibogae* and *Rhizophysa filiformis* (Fig. \@ref(siphonophores)A) instead move through the water column using repeated contraction and relaxation of the stem, and use a gas-filled float and, in the case of *B. sibogae*, modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore].  Nectophores are also present within the gonodendra (reproductive structures) of cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960]. It is not clear whether the nectophores found within the siphosome of the cystonects are homologous to the nectophores borne on the nectosome of codonophorans. Similarly, the homology of the special nectophore associated with gonophores of the calycophoran *Stephanophyes superba* is also unclear [@chun1891canarischen]. In this study, we only consider the evolution of the nectosome, and not the presence/absence of nectophores. The present analyses, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig. \@ref(fig:nectosome-origin), \@ref(fig:nectosome-BI)).

@@

Gastrozooids (specialized feeding polyps) have a single tentacle attached to the base of the zooid that is used for prey capture (with the exception of *Physalia physalis*, which has separate zooids for feeding and prey capture). As in other cnidarians, stinging capsules, arranged in dense batteries of nematocysts, play a critical role in prey capture. In many siphonophore species these batteries are found in side branches of the tentacle, termed tentilla (Fig. \@ref(colony)A).  Outside of Siphonophora, most hydrozoans bear simple tentacles without side branches. It is still an open question whether the common ancestor of Siphonophora had tentilla. The only siphonophores species regarded as lacking tentilla are *P. physalis*, *Apolemia* spp. (Fig. \@ref(siphonophores)H), and *Bathyphysa conifera* (Fig. \@ref(siphonophores)B). Since *B. conifera* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we assume this is a case of secondary loss. When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support a common ancestor bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig. \@ref(fig:tentilla), \@ref(fig:tentilla-BI)). However, this leaves a 30% support for a simple-tentacled common ancestor followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and non-apolemiid codonophorans.

@@

Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Fig. \@ref(traitdist), \@ref(fig:median-cont), \@ref(fig:depth-BI)) [@mapstone2014global; @mackie1987siphonophore]. The depth distribution of siphonophore populations is not always static, as some species are known to be vertical migrators, although this is within a relatively narrow depth range (<100m) [@pugh1984diel]. Some species such as *Nanomia bijuga* exhibit synchronous diel migration patterns [@barham1966deep]. Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig. \@ref(fig:median-cont) and \@ref(fig:depth-BI)), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck). This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic and bathypelagic waters. There was only a single transition to benthic lifestyle on the branch of Rhodaliidae, and a single transition to a pleustonic lifestyle on the branch of *P. physalis*. There is evidence that habitat depth is conserved within some clades, with the exception of Calycophorae which have diversified across the water column (Fig. \@ref(fig:median-cont), \@ref(fig:depth-BI)). Under the ML topology, depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow-living species) and Group B (deep-dwelling species) (Fig. \@ref(fig:median-cont)), while under the BI-CAT topology, a mesopelagic common ancestor is predicted, with a transition to epipelagic waters in Clade A (Fig. \@ref(fig:depth-BI)); however several shallow-living species that likely belong in Group B were not included in this analysis. The present sampling is also not sufficient to capture significant variation in depth distributions between closely related species. Previous studies have shown that many species that are collected at the same locality are found to occupy discrete, largely non-overlapping depth distributions, including between species that are closely related [@pugh1974vertical]. This suggests that vertical habitat use is more labile than it appears and may be an important mechanism in siphonophore ecology. The observed variation in depth distribution could be attributed to any of the correlated environmental variables (i.e. temperature, chlorophyll, oxygen). Temperature has been hypothesized to impose physiological limits to the dispersal of some clausophyid siphonophores [@grossmann2015diversity]. Since we only have oceanographic data from the Monterey Bay region, our analyses cannot disentangle the effects of these different variables on the vertical distributions.

@@

This work was supported by the National Science Foundation (DEB-1256695 and the Waterman Award). CM was also supported in part by a RI-EPSCoR Fellowship, NSF EPS-1004057. Sequencing at the Brown Genomics Core facility was supported in part by NIH P30RR031153 and NSF EPSCoR EPS-1004057. Data transfer was supported by NSF RII-C2 EPS-1005789. Analyses were conducted with computational resources and services at the Center for Computation and Visualization at Brown University, supported in part by the NSF EPSCoR EPS-1004057 and the State of Rhode Island. SOWHAT analyses were carried out on the Odyssey cluster supported by the FAS Division of Science, Research Computing Group at Harvard University -- we thank Cassandra Extavour for use of the Harvard cluster. Analyses were also conducted on the Hoffman2 cluster of the Institute for Digital Research and Education of the University of California, Los Angeles. We thank Rob Sherlock for providing the *Bathyphysa conifera* photograph. We also thank Zack Lewis for help with RNA extractions and the MBARI crews and ROV pilots for collection of the specimens.

@@

```{r supplementary-trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution in IQTree. We used ModelFinder [@kalyaanamoorthy2017modelfinder] to identify which examined model as the best relative model fit. We also present results for the commonly used WAG and GTR models. Unlabeled nodes have support >0.99. A. JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories (the model identified by ModelFinder as having the best fit). Log likelihood:-8113694.922; AIC:16227609.9565 .  B. GTR + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories. Log likelihood: -8133157.335; AIC:16266530.7277. C. WAG + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories. Log likelihood: -8156043.772; AIC score: 16312303.6120 "}

>>>>

Siphonophores remain poorly known, in large part because they are fragile and difficult to collect. They have, however, been of great interest for more than 150 years due to their unique structure and development [@mapstone2014global;@mackie1987siphonophore]. Like many other cnidarians, they are colonial: they grow by incomplete asexual reproduction. Each colony arises from a single embryo that forms the protozooid, the first body. One or two growth zones (Fig. \@ref(colony)) then arise and asexually produce other genetically identical zooids that remain attached [@carre1969etude; @carre1991complete; @carre1967; @carre1993ordre]. In some species additional zooids are added outside the growth zone along the siphosomal stem [@siebert2013re]. These zooids are each homologous to a solitary animal, but are physiologically integrated [@totton1965synopsis; @mackie1987siphonophore; @dunn2006evolution]. Siphonophores differ significantly from other colonial animals in their colony structure and development -- their zooids are highly functionally specialized and arranged in precise, repeating, species-specific patterns [@Beklemishev1969;@cartwright2010character]. Zooids are specialized for a range of functions, including feeding, reproduction, or swimming (Fig. \@ref(colony)) [@dunn2006evolution].

@@

Understanding the unique ecology, morphology, and development of siphonophores requires a well-resolved phylogeny of the group. The relationship of siphonophores to other hydrozoans has been difficult to determine [@cartwright2010character; @Cartwright2008kl; @zapata2015phylogenomic; @kayal2013cnidarian; @kayal2015phylogenetic; @kayal2018phylogenomics], but there has been progress on their internal relationships. A phylogeny [@Dunn:2005dy] based on two genes (16S, 18S) from 52 siphonophore taxa addressed several long standing questions about siphonophore biology, including the relationships of the three historically recognised groups, Cystonectae, Physonectae, and Calycophorae. Cystonectae was found to be sister to all other siphonophores, while Calycophorae were nested within "Physonectae". The name Codonophora was given to this clade of "Physonectae" and Calycophorae [@Dunn:2005dy].

Major questions remained after this early work, though. In particular, there was little support for important deep relationships within Codonophora. Understanding these relationships is key to resolving the evolution of several traits of importance, including sexual systems (monoecy versus dioecy) and the gain and loss of particular zooids, such as palpons (Fig. \@ref(colony)). Here we present a broadly sampled phylogenetic analysis of Siphonophora that considers transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded due to poor sampling). Using 1,423 genes, we find strong support for many relationships found in the earlier phylogeny [@Dunn:2005dy], and also provide new resolution for key relationships that were unresolved in that previous study. Using this phylogeny, we reconstruct the evolutionary history of characters central to the unique biology of siphonophores, including zooid type, life history traits, and vertical habitat use.

@@

All scripts for the analyses are available in a git repository at [https://github.com/caseywdunn/siphonophore_phylogeny_2017](https://github.com/caseywdunn/siphonophore_phylogeny_2017). The most recent commit at the time of the analysis presented here was `r system("git log | head -n 1", intern=TRUE) %>% str_replace("commit ", "")` with tag "paper_v2".

@@

We used ModelFinder [@kalyaanamoorthy2017modelfinder], as implemented in IQTree v1.6.3 [@doi:10.1093/molbev/msu300], to assess relative model fit. ModelFinder selected JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories based on the Bayesian Information Criterion. To assess the robustness of our results, we conducted phylogenetic analyses using multiple software programs, methods (Maximum likelihood (ML) and Bayesian Inference (BI)), and models (including the model selected by ModelFinder and several other commonly used models). Maximum likelihood (ML) analyses were conducted with RAxML v8.2.0 [@stamatakis2006raxml] and IQTree v1.6.3 [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281]. Bayesian Inference (BI) were conducted with Phylobayes v. 1.7a-mpi [@lartillot2009phylobayes]. Sequence alignments, sampled and consensus trees, and voucher information are available in the git repository. Tree figures were rendered with ggtree [@Yu:2016fo].

RAxML ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution (Fig. \@ref(phylogeny)A). RAxML bootstrap values were estimated using `r length(bootstrap_trees)` replicates. IQTree ML analyses were run under JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories, the best model identified by ModelFinder, and the commonly used model GTR + Optimized base frequencies + Free rate model with 6 categories ((Fig. \@ref(fig:supplementary-trees))).

BI was conducted in phylobayes using two different models: fixed-state WAG+$\Gamma$ (Fig. \@ref(fig:BI-WAG)) and CAT-Poisson (Fig. \@ref(bayesian)). Eight chains were run under the CAT-Poisson model. Four chains were run udner WAG+$\Gamma$, these runs did not converge (maxdiff=1, meandiff=0.0130273). The CAT-Poisson runs did not converge (maxdiff=1, meandiff=0.0565898). Closer inspection revealed that chain1 and chain3 were stuck in local maxima with low likelihood relative to other chains after 1405 and 4695 iterations. These two chains were excluded from the analyses, and the results presented here are based on the remaining 6 CAT-Poisson chains (maxdiff=1, meandiff=0.0185032). Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs. This left `r length( phylobayes_trees )` trees in the CAT-Poisson posterior.

@@

Morphological character data used in trait mapping were obtained from the literature or direct observation of available voucher material. Depth distribution data was queried from the MBARI VARS database ( http://www.mbari.org/products/research-software/video-annotation-and-reference-system-vars/) [@schlining2006mbari]. We used stochastic character mapping to infer the most probable evolution of traits on the tree in R using the `phytools` package [@MEE3:MEE3169; @huelsenbeck2003stochastic]. Subsequent analyses were conducted in R and integrated into this manuscript with the `knitr` package. See Supplementary Information for R package version numbers.

@@

The phylogenetic relationships recovered in this study received strong support across analysis methods (Fig. \@ref(phylogeny)A), with a couple of localized exceptions (Fig. \@ref(phylogeny)B and \@ref(bayesian)). All of the ML analyses were congruent with each other, regardless of model and software used (Fig. \@ref(fig:supplementary-trees)). These ML results were also congruent with the Phylobayes BI WAG+$\Gamma$ analyses (Fig. \@ref(fig:BI-WAG)). The Phylobayes BI CAT-Poisson result (Fig. \@ref(bayesian)), however, had a strongly supported topology that differed (Fig. \@ref(phylogeny)B) from the ML topology in localized regions as described below. The fact that the Phylobayes BI WAG+$\Gamma$ is consistent with the WAG (and other) ML analyses suggests that the different topology recovered in the Phylobayes BI CAT-Poisson analyses is due to the different model rather than different software or methods. Here we take the conservative approach of considering relationships that differ between the Phylobayes BI CAT-Poisson analyses and other analyses to be unresolved.

@@

In ML analyses and BI WAG analyses, Euphysonectae consists of two reciprocally monophyletic groups that we here provisionally refer to as Clade A and Group B (Fig. \@ref(phylogeny)A). In BI CAT-Poisson analyses, Group B is paraphyletic (Fig. \@ref(phylogeny)B). The presence of an involucrum, a fold around the base of the cnidoband [@totton1965synopsis], is a potential synapomorphy for Clade A. Species of Clade A also have a descending mantle canal within the nectophores (Fig. \@ref(fig:desc-mantle), \@ref(fig:mantle-BI)), a structure that is also present in some calycophorans. Members of Clade A are also monoecious (Fig. \@ref(reconstruct)). There is not a clear synapomorphy for Group B. Within Group B there is high support for the placement of *Erenna richardi* in ML analyses and BI WAG (Fig. \@ref(phylogeny) and \@ref(fig:BI-WAG)), but it is placed as sister to Clade A in BI CAT-Poisson analyses (Fig. \@ref(phylogeny)B). More taxon sampling will be required to determine the relationship of species within this group.

@@

Nectophores (Fig. \@ref(colony)) are non-reproductive propulsive medusae. In Codonophora, the nectophores are localized to a region known as the nectosome (Fig. \@ref(colony)B), which has its own growth zone, and they are used for coordinated colony-level swimming. Planktonic cystonects like *Bathyphysa sibogae* and *Rhizophysa filiformis* (Fig. \@ref(siphonophores)A) instead move through the water column using repeated contraction and relaxation of the stem, and in the case of *B. sibogae*, use modified flattened gastrozooids with wings (called ptera) to increase surface area and prevent colony sinking [@biggs1976siphonophore]. Nectophores are also present within the gonodendra (reproductive structures) of cystonects, and are thought to propel the gonodendra when they detach from the colony [@totton1965synopsis;@Totton1960]. It is not clear whether the nectophores found within the siphosome of the cystonects are homologous to the nectophores borne on the nectosome of codonophorans. Similarly, the homology of the special nectophore associated with gonophores of the calycophoran *Stephanophyes superba* is also unclear [@chun1891canarischen]. In this study, we only consider the evolution of the nectosome, and not the presence/absence of nectophores. The present analyses, as well as the analyses of @Dunn:2005dy, are consistent with a single origin of the nectosome (Fig. \@ref(fig:nectosome-origin), \@ref(fig:nectosome-BI)).

@@

Gastrozooids (specialized feeding polyps) have a single tentacle attached to the base of the zooid that is used for prey capture (with the exception of *Physalia physalis*, which has separate zooids for feeding and prey capture, and rhodalids, where some tentacles are used to anchor to the substrate and do not participate in feeding). As in other cnidarians, stinging capsules, arranged in dense batteries of nematocysts, play a critical role in prey capture. In many siphonophore species these batteries are found in side branches of the tentacle, termed tentilla (Fig. \@ref(colony)A). Outside of Siphonophora, most hydrozoans bear simple tentacles without side branches. It is still an open question whether the common ancestor of Siphonophora had tentilla. The only siphonophores species regarded as lacking tentilla are *P. physalis*, *Apolemia* spp. (Fig. \@ref(siphonophores)H), and *Bathyphysa conifera* (Fig. \@ref(siphonophores)B). Since *B. conifera* is the only member of the *Rhizophysidae* (and of the *Bathyphysa* genus) lacking tentilla, we assume this is a case of secondary loss. When we reconstruct the evolution of this character on the current phylogeny, 70% of simulations support a common ancestor bearing tentilla, with two independent losses leading to *Physalia* and *Apolemia* (Fig. \@ref(fig:tentilla), \@ref(fig:tentilla-BI)). However, this leaves a 30% support for a simple-tentacled common ancestor followed by 2 independent gains of tentilla in the branches leading to *Rhizophysidae* and non-apolemiid codonophorans.

@@

Siphonophores are abundant predators in the pelagic realm, ranging from the surface (*Physalia physalis*) to bathypelagic depths (Fig. \@ref(traitdist), \@ref(fig:median-cont), \@ref(fig:depth-BI)) [@mapstone2014global; @mackie1987siphonophore]. The depth distribution of siphonophore populations is not always static, as some species are known to be vertical migrators, although this is within a relatively narrow depth range (<100m) [@pugh1984diel]. Some species such as *Nanomia bijuga* exhibit synchronous diel migration patterns [@barham1966deep]. Using the present phylogeny, we reconstructed the median depth changes along the phylogeny under a Brownian Motion model (Fig. \@ref(fig:median-cont) and \@ref(fig:depth-BI)), which had the strongest AICc support (compared to non-phylogenetic distributions, and to Ohrnstein-Uhlenbeck). This model indicates a mesopelagic most recent common ancestor, with several independent transition events to epipelagic and bathypelagic waters. There was only a single transition to benthic lifestyle on the branch of Rhodaliidae, and a single transition to a pleustonic lifestyle on the branch of *P. physalis*. There is evidence that habitat depth is conserved within some clades, with the exception of Calycophorae which have diversified across the water column (Fig. \@ref(fig:median-cont), \@ref(fig:depth-BI)). Under the ML topology, depth appears to be phylogenetically conserved in Euphysonectae after the split between Clade A (shallow-living species) and Group B (deep-dwelling species) (Fig. \@ref(fig:median-cont)), while under the BI-CAT topology, a mesopelagic common ancestor is predicted, with a transition to epipelagic waters in Clade A (Fig. \@ref(fig:depth-BI)); however several shallow-living species that likely belong in Group B were not included in this analysis. The present sampling is also not sufficient to capture significant variation in depth distributions between closely related species. Previous studies have shown that many species that are collected at the same locality are found to occupy discrete, largely non-overlapping depth distributions, including between species that are closely related [@pugh1974vertical]. This suggests that vertical habitat use is more labile than it appears and may be an important mechanism in siphonophore ecology. The observed variation in depth distribution could be attributed to any of the correlated environmental variables (i.e. temperature, chlorophyll, oxygen). Temperature has been hypothesized to impose physiological limits to the dispersal of some clausophyid siphonophores [@grossmann2015diversity]. Since most of our specimens were sampled only in the Monterey Bay region, our analyses of the local oceanographic and depth distribution data cannot disentangle the effects of these different variables on the vertical distributions.

@@

This work was supported by the National Science Foundation (DEB-1256695, the Waterman Award, and DEB-1701272). CM was also supported in part by a RI-EPSCoR Fellowship, NSF EPS-1004057. Sequencing at the Brown Genomics Core facility was supported in part by NIH P30RR031153 and NSF EPSCoR EPS-1004057. Data transfer was supported by NSF RII-C2 EPS-1005789. Analyses were conducted with computational resources and services at the Center for Computation and Visualization at Brown University, supported in part by the NSF EPSCoR EPS-1004057 and the State of Rhode Island. SOWHAT analyses were carried out on the Odyssey cluster supported by the FAS Division of Science, Research Computing Group at Harvard University -- we thank Cassandra Extavour for use of the Harvard cluster. Analyses were also conducted on the Hoffman2 cluster of the Institute for Digital Research and Education of the University of California, Los Angeles. We thank Rob Sherlock for providing the *Bathyphysa conifera* photograph. We also thank Zack Lewis for help with RNA extractions and the MBARI crews and ROV pilots for collection of the specimens.

@@

```{r supplementary-trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution in IQTree. We used ModelFinder [@kalyaanamoorthy2017modelfinder] to identify the model with the best relative model fit. We also present results for the commonly used WAG and GTR models. Unlabeled nodes have support >0.99. A. JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories (the model identified by ModelFinder as having the best fit). Log likelihood:-8113694.922; AIC:16227609.9565 .  B. GTR + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories. Log likelihood: -8133157.335; AIC:16266530.7277. C. WAG + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories. Log likelihood: -8156043.772; AIC score: 16312303.6120 "}
